Synthesizing Quotient Lenses

### 2018-03-14 papers/QuotientLenses/main.tex
Nits

Recent work~\cite{optician}\saz{removed ``we'' here to maintain the double-blind
  review standard.} developed a tool called Optician to synthesize the
class of {\em bijective lenses} that can be derived in the lens programming
language {\em Boomerang}~\cite{boomerang}. Each lens $\ell : S \Leftrightarrow T$
that Optician can synthesize specifies a bijection from the language of a
source regular expression $S$ to a target language of a target regular
expression $T$: \begin{equation}\label{bijectivelenslaws} \ell.\get \;
(\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s) = s
\end{equation}
Optician makes programming bijective lenses faster and easier, and though
many useful lenses are not bijective in nature, they do have a bijective core
surrounded by structured transformations. Hence, Optician can be used as a
component in a system that synthesizes more complex lenses.

This paper applies this idea to the problem of
synthesizing {\em quotient lenses}~\cite{quotientlenses}.
Quotient lenses are lenses in which the lens laws are
loosened so that they hold modulo an equivalence relation on the source and
target data respectively; in this paper we are concerned with {\em bijective
quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
and target data respectively:
\begin{equation}\label{quotientlenslaws}
\ell.\get \; (\ell.\lput \; t) \equiv_T t \text{, and } \ell.\lput \; (\ell.\get
\; s) \equiv_S s
\end{equation}
Quotient lenses are useful in situations where a programmer wishes for the
transformation defined by a lens to have the same behavior on data that differ
only in inessential details. For instance, a programmer may wish to ``quotient
away'' the number of white space characters between data items, the
capitalization of various strings, the sequence of fields in a record, or the
order of items in a list.

@@

For example, consider the following QRE for writing author names:
%
\begin{lstlisting}
let wsp_sp = collapse wsp $\to$ " "
let comma_name = first_name . "," . wsp_sp . last_name
\end{lstlisting}
%
In this example, \lstinline{wsp} is an existing regular expression for a
nonempty sequence of whitespace characters, and \lstinline{first_name} and
\lstinline{last_name} are existing regular expressions for first and last names.
The QRE \lstinline{wsp_sp} is a QRE with the same underlying language as
\lstinline{wsp}, but with a single canonical representative: \lstinline{" "}.
This is then used to build up \lstinline{comma_name}, a QRE with an
underlying language of two names, separated by a command and whitespace, where
the names with a single space between them are canonical.

>>>>

Recent work~\cite{optician}\saz{removed ``we'' here to maintain the double-blind
  review standard.} developed a tool called Optician to synthesize the
class of {\em bijective lenses} that can be derived in the lens programming
language {\em Boomerang}~\cite{boomerang}\bcp{All of them?  I'm not certain
  that's accurate.}. Each lens $\ell : S \Leftrightarrow T$
that Optician can synthesize specifies a bijection from the language of a
source regular expression $S$ to a target language of a target regular
expression $T$: \begin{equation}\label{bijectivelenslaws} \ell.\get \;
(\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s) = s
\end{equation}
Optician makes programming bijective lenses faster and easier, and though
many useful lenses are not bijective in nature, they do have a bijective core
surrounded by structured transformations. Hence, Optician can be used as a
component in a system that synthesizes more complex lenses.

This paper applies this idea to the problem of
synthesizing {\em quotient lenses}~\cite{quotientlenses}.
Quotient lenses are lenses in which the lens laws are
loosened so that they hold modulo an equivalence relation on the source and
target data respectively; in this paper we are concerned with {\em bijective
quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
and target data respectively:\bcp{Not too convinced about the colors.  The
  fact that lens expressions mix black and red is distracting -- my eyes
  can't easily tell which bits of a larger expression are lenses.}
\begin{equation}\label{quotientlenslaws}
\ell.\get \; (\ell.\lput \; t) \equiv_T t \text{, and } \ell.\lput \; (\ell.\get
\; s) \equiv_S s
\end{equation}
Quotient lenses are useful in situations where a programmer wishes for the
transformation defined by a lens to have the same behavior on data that differ
only in inessential details. For instance, a programmer may wish to ``quotient
away'' the number of white space characters between data items, the
capitalization of various strings, the sequence of fields in a record, or the
order of items in a list.

@@

For example, consider the following QRE for writing author
names:\bcp{Solomon: Someone noted on their copy of the intro from
  class yesterday that this example is backwards.  Did you go over the
  marked-up copies that I gave you?}
%
\begin{lstlisting}
let wsp_sp = collapse wsp $\to$ " "
let comma_name = first_name . "," . wsp_sp . last_name
\end{lstlisting}
%
In this example, \lstinline{wsp} is an existing regular expression for a
nonempty sequence of whitespace characters, and \lstinline{first_name} and
\lstinline{last_name} are existing regular expressions for first and last names.
The QRE \lstinline{wsp_sp} is a QRE with the same underlying language as
\lstinline{wsp}, but with a single canonical representative: \lstinline{" "}.
This is then used to build up \lstinline{comma_name}, a QRE with an
underlying language of two names, separated by a command and whitespace, where
the names with a single space between them are canonical.
